---
title: "caa_multi_GEX"
author: "Chenxin Li"
date: "2023-10-20"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Package
```{r}
library(tidyverse)
library(Seurat) 
library(Signac)
library(readxl)

library(RColorBrewer)
library(viridis)
library(rcartocolor)

library(patchwork)
```

# Read starsolo output 
```{r}
BG_raw <- Read10X("../Results/CAA_BGGeneFull/raw/")
BI_raw <- Read10X("../Results/CAA_BIGeneFull/raw/")
```

```{r}
BG <- CreateSeuratObject(counts = BG_raw, min.cells = 3, min.features = 200,
                         project = "BG")

BI <- CreateSeuratObject(counts = BI_raw, min.cells = 3, min.features = 200,
                         project = "BI")
```


```{r}
BG
BI
```

# Filtering
```{r}
blank <- data.frame(
  x = 1:10,
  y = 1:10
) %>% 
  ggplot(aes(x = x, y = y)) +
  theme_void()
```

## BG
```{r}
BG_scatter <- BG@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 200, color = "red2") +
  geom_vline(xintercept = 4000, color = "red2") +
  geom_hline(yintercept = 200, color = "blue1") +
  geom_hline(yintercept = 4000, color = "blue1") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

BG_hist1 <- BG@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 200, color = "red2") +
  geom_vline(xintercept = 4000, color = "red2") +
  scale_x_log10() +
  theme_classic()

BG_hist2 <- BG@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 200, color = "blue1") +
  geom_vline(xintercept = 4000, color = "blue1") +
  scale_x_log10() +
  theme_classic() +
  coord_flip()

wrap_plots(
  BG_hist1, blank,
  BG_scatter, BG_hist2,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.2), 
  heights = c(0.2, 1)
)

ggsave("../Results/R_output/BG_filtering2.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_output/BG_filtering2.png", height = 4, width = 5, bg = "white")
```

```{r}
BG_sub <- subset(BG, subset = nFeature_RNA > 200 &
                   nFeature_RNA < 4000 &
                   nCount_RNA < 4000 &
                   nCount_RNA > 200)

BG_sub
```


## BI 
```{r}
BI_scatter <- BI@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 200, color = "red2") +
  geom_vline(xintercept = 4000, color = "red2") +
  geom_hline(yintercept = 200, color = "blue1") +
  geom_hline(yintercept = 4000, color = "blue1") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

BI_hist1 <- BI@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 200, color = "red2") +
  geom_vline(xintercept = 4000, color = "red2") +
  scale_x_log10() +
  theme_classic()

BI_hist2 <- BI@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 200, color = "blue1") +
  geom_vline(xintercept = 4000, color = "blue1") +
  scale_x_log10() +
  theme_classic() +
  coord_flip()

wrap_plots(
  BI_hist1, blank,
  BI_scatter, BI_hist2,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.2), 
  heights = c(0.2, 1)
)

ggsave("../Results/R_output/BI_filtering2.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_output/BI_filtering2.png", height = 4, width = 5, bg = "white")
```

```{r}
BI_sub <- subset(BI, subset = nFeature_RNA > 200 &
                   nFeature_RNA < 4000 &
                   nCount_RNA < 4000 &
                   nCount_RNA > 200)

BI_sub
```

# Summaries
```{r}
lib_summary <- data.frame(
  ID = c("BG", "BI"),
  cells = c(ncol(BG_sub), 
            ncol(BI_sub)),
  median_UMI = c(summary(BG_sub$nCount_RNA)[3],
                 summary(BI_sub$nCount_RNA)[3]),
  median_genes = c(summary(BG_sub$nFeature_RNA)[3],
                   summary(BI_sub$nFeature_RNA)[3])
)

lib_summary

write_excel_csv(lib_summary, "../Results/R_output/leaf_gex_lib_summary_stringent.csv")
```
# Normalize, find variable features 
```{r}
BG_sub <- NormalizeData(BG_sub, normalization.method = "LogNormalize", scale.factor = 10000)
BI_sub <- NormalizeData(BI_sub, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
BG_sub <- FindVariableFeatures(BG_sub, selection.method = "vst", nfeatures = 3000)
BI_sub <- FindVariableFeatures(BI_sub, selection.method = "vst", nfeatures = 3000)
```
# Integrate
```{r}
RNA.list <- list(BG_sub, BI_sub)
features <- SelectIntegrationFeatures(object.list = RNA.list, nfeatures = 3000)
anchors <- FindIntegrationAnchors(object.list = RNA.list, anchor.features = features) 
CAA_L2 <- IntegrateData(anchorset = anchors)
```
# Scale data
```{r}
CAA_L2 <- ScaleData(CAA_L2, verbose = F)

CAA_L2@meta.data$run <- CAA_L2@meta.data$orig.ident

CAA_L2@meta.data %>% 
  group_by(run) %>% 
  count()
```

# PCA
```{r}
CAA_L2 <- RunPCA(CAA_L2, verbose = FALSE, npcs = 50) # run PCA

DimPlot(CAA_L2, reduction = "pca", cols = brewer.pal(8, "Set2")[c(1:3)],
        split.by = "run", group.by = "run") +
theme(
  legend.position = "none"
)
```

# UMAP
```{r}
CAA_L2 <- RunUMAP(CAA_L2, dims = 1:30, 
                     min.dist = 0.001, repulsion.strength = 1, 
                     n.neighbors = 30, spread = 1
                    ) 

CAA_L2 <- FindNeighbors(CAA_L2, reduction = "pca", dims = 1:50)
CAA_L2 <- FindClusters(CAA_L2, resolution = 0.5)
```

```{r}
UMAP_RNA <- DimPlot(CAA_L2, reduction = "umap", 
                 label = T, label.size = 5, repel = T) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none"
  )

UMAP_RNA

ggsave("../Results/R_output/UMAP_caa1_stringent.svg", height = 3, width = 3, bg = "white")
ggsave("../Results/R_output/UMAP_caa1_stringent.png", height = 3, width = 3, bg = "white")
```

## Check replicates 
```{r}
DimPlot(CAA_L2, reduction = "umap", 
        label = T, label.size = 5, repel = T, split.by = "run") +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) +
  ggtitle("Grouped by replicates\n")  
  

ggsave("../Results/R_output/UMAP_RNA_by_rep_strigent.svg", height = 3, width = 8, bg = "white")
ggsave("../Results/R_output/UMAP_RNA_by_rep_stringent.png", height = 3, width = 8, bg = "white")
```
# De novo markers 
```{r}
all_makers <- FindAllMarkers(CAA_L2, only.pos = T, min.pct = 0.05)
all_makers <- all_makers %>% 
  filter(p_val_adj < 0.05)

head(all_makers)
```

## Pull reprentative loci 
```{r}
repr_locus <-  read_delim("../Data/caac_v4.hc_gene_models.repr.gtf", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE, skip = 1)


repr_locus %>% 
  filter(str_detect(X9, "nbisL1-mrna-18701"))

repr_locus_nice <- repr_locus %>% 
  filter(X3 == "gene") %>% 
  select(X1, X4, X5, X9) %>% 
  separate(X9, c("gene_id", "ID", "Name"), sep = "; ") %>% 
  mutate(ID2 = str_extract(ID, "nbisL1-mrna-\\d+")) %>% 
  mutate(v4_ID = str_sub(Name, start = 7, end = -3)) %>%
  select(-gene_id, -ID, -Name) %>% 
  mutate(v4_locus = str_remove(v4_ID, "\\.\\d+$"))
  
  
head(repr_locus_nice)
```

## Pull functional annotation 
```{r}
func_anno <- read_delim("../Data/caac_v4_anno/caac_v4.functional_annotation.txt", 
                        delim = "\t", col_names = F)

func_anno <- func_anno %>% 
  rename(v4_ID = X1)
head(func_anno)
```

## Pull convertion between v3 and v4 IDs 
```{r}
v3_v4_convertion <- read_csv("../Results/v3_v4_convertion.csv")

v3_v4_convertion_nice <- v3_v4_convertion %>% 
  select(qseqid, sseqid) %>% 
  rename(v3_ID = qseqid) %>% 
  rename(v4_ID = sseqid) %>% 
  mutate(v3_ID = str_remove(v3_ID, "\\.\\d+$")) %>% 
  mutate(v4_locus = str_remove(v4_ID, "\\.\\d+$")) %>% 
  select(-v4_ID)

head(v3_v4_convertion_nice)
```

```{r}
gene_ID_table <- repr_locus_nice %>% 
  inner_join(v3_v4_convertion_nice, by = "v4_locus") %>% 
  inner_join(func_anno, by = "v4_ID")

write_excel_csv(gene_ID_table, "../Results/gene_ID_table.csv")
```

```{r}
all_makers_nice <- all_makers %>% 
  left_join(repr_locus_nice, by = c("gene"="ID2")) %>% 
  left_join(func_anno, by = c("v4_ID"))

head(all_makers_nice)

write_excel_csv(all_makers_nice, "../Results/R_output/leaf_de_novo_markers.csv") 
```


# Check marker genes
```{r}
leaf_marker <- read_csv("../Data/caa.leaf_marker.csv")
head(leaf_marker)

leaf_marker2 <- leaf_marker %>% 
  select(X1, X2, order, symbol) %>% 
  rename(gene_ID = X2) %>% 
  distinct(gene_ID, .keep_all = T) %>% 
  rename(v3_ID = gene_ID) %>% 
  mutate(v3_ID = str_remove(v3_ID, "\\.\\d+$")) %>%
  inner_join(v3_v4_convertion_nice, by = "v3_ID") %>% 
  inner_join(repr_locus_nice, by = "v4_locus") %>% 
  filter(symbol != "SUC2") %>% 
  select(order, symbol, v4_locus, ID2, v4_ID) %>% 
  add_case(order = 13, symbol = "STR1", ID2 = "nbisL1-mrna-17182", v4_ID = "Caacu.v4.8G010020.1") %>% 
  add_case(order = 14, symbol = "STR2", ID2 = "nbisL1-mrna-17180", v4_ID = "Caacu.v4.8G010000.1") %>% 
  add_case(order = 15, symbol = "histone", ID2 = "nbisL1-mrna-26163", v4_ID = "Caacu.v4.14G009410.1") %>% 
  add_case(order = 16, symbol = "histone", ID2 = "nbisL1-mrna-13277", v4_ID = "Caacu.v4.6G008330.1") %>% 
  add_case(order = 17, symbol = "cyclin", ID2 = "nbisL1-mrna-19537", v4_ID = "Caacu.v4.9G014530.1") %>% 
  add_case(order = 18, symbol = "cyclin", ID2 = "nbisL1-mrna-14705", v4_ID = "Caacu.v4.7G001510.2") 

head(leaf_marker2, 20)
```

```{r}
levels(CAA_L2) <- c(
  0, 1, 10, 12, 15,      # Meso
  3, 5, 11, 13,  # Ep
          # GC
  2, 4, 14, 17, # V
  8, # STR
  7, # S-phase
  6, # Mitosis  
  9, 16   # Un
)
```

```{r}
DefaultAssay(CAA_L2) <- "RNA"
marker_at_cluster <- DotPlot(CAA_L2, features = unique(rev(leaf_marker2$ID2)))

leaf_marker_curated <- marker_at_cluster$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene_ID = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  left_join(leaf_marker2, by = c("features.plot"="ID2")) %>%
  mutate(symbol = reorder(symbol, -order)) %>% 
  mutate(v4_ID = reorder(v4_ID, -order)) %>% 
  mutate(pct.exp_clip = case_when(
    pct.exp >= 30 ~ 30,
    T ~ pct.exp
  ))
 
marker_heat <- leaf_marker_curated %>% 
  ggplot(aes(y = v4_ID, x = id)) +
  geom_point(aes(size = pct.exp_clip, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Average Exp.",
       size = "% Expressed",
       title = "leaf marker genes") +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5),
        legend.position = "top",
        legend.key.width = unit(0.7, "lines"),
        legend.key.height = unit(0.7, "lines"),
        legend.box = "horizontal"
        )

marker_heat
```
## Assign cell types 
 0, 1, 10, 12, 15,      # Meso
  3, 5, 11, 13,  # Ep
          # GC
  2, 4, 14, 17, # V
  8, # STR
  7, # S-phase
  6, # Mitosis  
  9, 16   # Un
  
```{r}
caa_cell_types <- data.frame(
  id = 0:17
) %>% 
  mutate(id = factor(id, levels = c(
  0, 1, 10, 12, 15, # Meso
  3, 5, 11, 13,  # Ep
          # GC
  2, 4, 14,  17, # V
  8, # STR
  7, # S-phase 
  6, # mitosis 
  9, 16 # Un
  ))) %>%
  mutate(cell_type = case_when(
   id %in% c(0, 1, 10, 12, 15)
     ~ "Mesophyll",
  id == 3 |
     id == 5 |
     id == 11 |
    id == 13 ~ "Epidermis",
  id == 2 |
     id == 4 |
     id == 14 |
    id == 17 ~ "Vasculature",
    id == 8 ~ "STR+",
    id == 6 ~ "Mitosis",
  id == 7 ~ "S phase",
  T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll",
                              "Epidermis", 
                             # "Guard cells", 
                              "Vasculature",
                             "STR+",
                             "S phase",
                              "Mitosis",
                              "Unassigned"
                            )))

caa_cell_types
```

## Percentage of cell types?
```{r}
head(CAA_L2@meta.data)
CAA_L2@meta.data <- CAA_L2@meta.data %>%
  mutate(cell_type = case_when(
   seurat_clusters %in% c(0, 1, 10, 12, 15)
     ~ "Mesophyll",
  seurat_clusters == 3 |
     seurat_clusters == 5 |
     seurat_clusters == 11 |
    seurat_clusters == 13 ~ "Epidermis",
  seurat_clusters == 2 |
     seurat_clusters == 4 |
     seurat_clusters == 14 |
    seurat_clusters == 17 ~ "Vasculature",
    seurat_clusters == 8 ~ "STR+",
    seurat_clusters == 6 ~ "Mitosis",
  seurat_clusters == 7 ~ "S phase",
  T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll",
                              "Epidermis", 
                              "Vasculature",
                             "STR+",
                             "Mitosis",
                             "S phase",
                              "Unassigned"
                            )))

Cell_type_summary <- CAA_L2@meta.data %>% 
  group_by(cell_type) %>% 
  count() %>% 
  mutate(total = ncol(CAA_L2)) %>% 
  mutate(percentage = n / total * 100)

Cell_type_summary

write_excel_csv(Cell_type_summary, "../Results/R_output/Cell_type_summary_stringent.csv")
```


```{r}
caa_assignment_graph <- caa_cell_types %>% 
  ggplot(aes(x = id, y = "")) +
  facet_wrap(~ cell_type, scales = "free", ncol = 2) +
  geom_point(size = 4, aes(color = id)) +
  geom_text(aes(label = id)) +
  labs(x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    text = element_text(color = "black", face = "bold"),
    axis.text.x = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )

caa_assignment_graph
```
 
```{r}
UMAP2 <- DimPlot(CAA_L2, reduction = "umap", 
                 label = T, label.size = 5, repel = T) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none"
  )

wrap_plots(UMAP2, caa_assignment_graph,
           nrow = 2, heights = c(1, 0.9))

ggsave("../Results/R_output/UMAP_2_stringent.svg", height = 4, width = 3.2, bg = "white")
ggsave("../Results/R_output/UMAP_2_stringnet.png", height = 4, width = 3.2, bg = "white")
```

```{r}
UMAP3 <- DimPlot(CAA_L2, reduction = "umap", group.by = "cell_type",  
                 label = T, label.size = 5, repel = T) + 
  scale_color_manual(values = c(brewer.pal(6, "Accent")[1:2],
                                brewer.pal(6, "Set2")[6],
                                brewer.pal(6, "Accent")[6],
                                brewer.pal(11, "BrBG")[2:1],
                                "grey80")) +
  labs(title = NULL) +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none"
  ) 

UMAP3

ggsave("../Results/R_output/UMAP_3_stringent.svg", height = 3.5, width = 4, bg = "white")
ggsave("../Results/R_output/UMAP_3_stringnet.png", height = 3.5, width = 4, bg = "white")
```


```{r}
caa_color_strip <- caa_cell_types %>% 
  ggplot(aes(x = id, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(6, "Accent")[1:2],
                                brewer.pal(6, "Set2")[6],
                                brewer.pal(6, "Accent")[6],
                                brewer.pal(11, "BrBG")[1:2],
                                "grey80")) +
  labs(fill = "Cell type") +
  guides(fill = guide_legend(nrow = 3)) +
  theme_void() +
  theme(
    legend.position = "bottom"
  )

caa_color_strip
```

```{r}
leaf_marker_symbol <- leaf_marker_curated %>% 
  filter(id == 0) %>% 
  ggplot(aes(x = "", y = v4_ID)) +
  # facet_grid(cell_type_order ~ ., scales = "free", space = "free") +
  geom_text(aes(label = symbol), fontface = "italic") +
  labs(x = NULL,
       y = NULL) +
  theme_classic() +
  theme(#text = element_text(color = "black"),
        axis.text = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        axis.line = element_blank())

leaf_marker_symbol
```

```{r}
wrap_plots(marker_heat, leaf_marker_symbol, caa_color_strip,
            design = c("AB
                       C#"), 
           heights = c(1, 0.03), 
           widths = c(1, 0.124)) 

ggsave("../Results/R_output/marker_genes_strignet.svg", height = 6, width = 7, bg = "white")
ggsave("../Results/R_output/marker_genes_strigent.png", height = 6, width = 7, bg = "white")
```

# Check pathway genes 
## Load known genes 
```{r}
known_genes <- read_csv("../Data/known_genes_nice_ordered.csv")

head(known_genes)
```


```{r}
known_genes_nice <- known_genes %>% 
  arrange(order2) %>% 
  filter(is.na(order2) ==F) %>% 
  dplyr::rename(v4_ID = sseqid) %>% 
  select(v4_ID, order2, symbol) %>% 
   rbind(
    data.frame(
      v4_ID = "Caacu.v4.10G004300.1",
      order2 = 10.5,
      symbol = "GPPS SSU"
    )
  ) %>% 
  inner_join(repr_locus_nice, by = "v4_ID") %>% 
  distinct(v4_ID, .keep_all = T) %>% 
  filter(symbol != "GPPS")

known_genes_nice
```

```{r}
write_excel_csv(known_genes_nice, "../Results/R_output/known_genes_nice_MEP.csv") 
```


```{r}
DefaultAssay(CAA_L2) <- "RNA"
known_genes_at_cluster <- DotPlot(CAA_L2, features = rev(known_genes_nice$ID2))
```
 
## Nicer dot plots 
```{r}
known_heat <- known_genes_at_cluster$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(ID2 = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  left_join(known_genes_nice, by = "ID2") %>%
  mutate(symbol = reorder(symbol, -order2)) %>% 
  mutate(pct.exp_clip = case_when(
    pct.exp >= 30 ~ 30,
    T ~ pct.exp
  )) %>% 
  filter(order2>=14) %>% 
  ggplot(aes(y = symbol, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Avg. Exp.",
       size = "% Exp.") +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.7, "lines"),
        legend.key.height = unit(0.7, "lines"),
        legend.position = "top",
        legend.box = "vertical",
        legend.spacing = unit(0, "lines"))

wrap_plots(known_heat, caa_color_strip + labs(fill = NULL),
           heights = c(1, 0.05), nrow = 2) 

ggsave("../Results/R_output/known_genes_strigent.svg", height = 4, width = 4.1, bg = "white")
ggsave("../Results/R_output/known_genes_strigent.png", height = 4, width = 4.1, bg = "white")
```


# Save object
```{r}
saveRDS(CAA_L2, "../Results/R_output/CAA_L2_strigent.Rds")
CAA_L2 <- readRDS("../Results/R_output/CAA_L2_strigent.Rds")
```

# Check stem TFs in leaf 
```{r}
BIS_clade <- repr_locus_nice %>% 
  filter(str_detect(v4_ID, "Caacu.v4.4G008460.4")) %>% 
  mutate(clade = "CrBIS\nclade") %>% 
    # inner_join(v3_v4_convertion_nice, by = "v3_ID") %>% 
    select(v4_locus, ID2, v4_ID, clade)

IDB_clade <- data.frame(
  v3_ID = c(
  "Caacu.2G012920",
  "Caacu.2G002990",
  "Caacu.7G015120") ) %>% 
  inner_join(v3_v4_convertion_nice, by = "v3_ID") %>% 
  left_join(repr_locus_nice, by = "v4_locus") %>% 
  select(v4_locus, ID2, v4_ID) %>% 
   mutate(clade = "CaTT8/CrIDB1\nclade")
  
IDM3_clade <- data.frame(
  v3_ID = c(
  "Caacu.1G002030", 
  "Caacu.1G020620" ,
  "Caacu.S014450",
  "Caacu.16G001190",
  "Caacu.4G003180",
  "Caacu.8G018100",
  "Caacu.12G012510",
  "Caacu.3G007010" ,
  "Caacu.9G000860",
  "Caacu.9G000870",
  "Caacu.8G018120",
  "Caacu.4G003170",
  "Caacu.12G012490",
  "Caacu.19G011850") ) %>% 
  inner_join(v3_v4_convertion_nice, by = "v3_ID") %>% 
  left_join(repr_locus_nice, by = "v4_locus") %>% 
  select(v4_locus, ID2, v4_ID) %>% 
  mutate(clade = "AtMYB123/CrIDM1\nclade")

IDM4_clade <- data.frame(
  v3_ID = c(
  "Caacu.20G000670",
  "Caacu.16G006210",
  "Caacu.2G015520") ) %>% 
  inner_join(v3_v4_convertion_nice, by = "v3_ID") %>% 
  left_join(repr_locus_nice, by = "v4_locus") %>% 
  select(v4_locus, ID2, v4_ID) %>% 
   mutate(clade = "AtMYB5/CrIDM4\nclade")
  
stem_TFs <- rbind(
  BIS_clade,
  IDB_clade,
  IDM3_clade,
  IDM4_clade
)

stem_TFs
```
```{r}
Leaf_myb123 <- FeaturePlot(CAA_L2, features = c("nbisL1-mrna-25208"),
            order = T) +
  scale_color_gradientn(colours = c("grey90", brewer.pal(9, "YlGnBu"))) +
  ggtitle("CaMYB123") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.7, "lines"))

# CaBIS not expressed in leaf 

leaf_caTT2 <- FeaturePlot(CAA_L2, features = c("nbisL1-mrna-15970"),
            order = T) +
  scale_color_gradientn(colours = c("grey90", brewer.pal(9, "YlGnBu"))) +
  ggtitle("CaTT2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.7, "lines"))

leaf_caMyb5 <- FeaturePlot(CAA_L2, features = c("nbisL1-mrna-29301"),
            order = T) +
  scale_color_gradientn(colours = c("grey90", brewer.pal(9, "YlGnBu"))) +
  ggtitle("CaMYB5") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.7, "lines"))

wrap_plots(Leaf_myb123, leaf_caTT2, leaf_caMyb5, nrow = 1)

ggsave("../Results/R_output/stem_TF_in_leaf_featureplot.svg", height = 3, width = 7)
ggsave("../Results/R_output/stem_TF_in_leaf_featureplot.png", height = 3, width = 7)
```

```{r}
DotPlot(CAA_L2, features = stem_TFs$ID2)$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(ID2 = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  left_join(stem_TFs, by = "ID2") %>%
  #mutate(symbol = reorder(symbol, -order2)) %>% 
  # mutate(pct.exp_clip = case_when(
  #   pct.exp >= 30 ~ 30,
  #   T ~ pct.exp
  # )) %>% 
  #filter(order2>=14) %>% 
  ggplot(aes(y = v4_ID, x = id)) +
  facet_grid(clade ~., scales = "free", space = "free") +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Avg. Exp.",
       size = "% Exp.") +
  theme_classic() +
  theme(text = element_text(color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5),
        legend.key.width = unit(0.7, "lines"),
        legend.key.height = unit(0.7, "lines"))

ggsave("../Results/R_output/stem_TF_in_leaf.svg", height = 5, width = 7)
ggsave("../Results/R_output/stem_TF_in_leaf.png", height = 5, width = 7)
```
Stem TFs are not co-expressed w/ biosynthetic genes in leaf. 
 
# Save gene exp matrix
```{r}
Leaf_gene_exp_matrix <- AverageExpression(CAA_L2, slot = "data")$RNA %>% 
  as.data.frame() %>% 
  mutate(ID2 = row.names(.)) %>% 
  left_join(repr_locus_nice, by = "ID2") %>% 
  select(v4_ID, 0:17) %>% 
  dplyr::rename(gene_ID = v4_ID)

head(Leaf_gene_exp_matrix)

write_excel_csv(Leaf_gene_exp_matrix, "../Results/R_output/leaf_gene_exp_matrix.csv")
```

 

