---
title: "MYB_bHLH_trees"
author: "Chenxin Li"
date: "2024-10-31"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(readxl)

library(RColorBrewer)
library(viridis)
library(rcartocolor)
library(svglite)

library(patchwork)

library(ggtree)
library(ggtreeExtra)
library(treeio)
library(ape)

library(Seurat)
```

# # MYBs
## Tree data 
```{r}
CAA_CRO_MYB_tree <- read.tree("../Results/Phylogeny/RAxML_bipartitions.myb_out")

CAA_CRO_MYB_tree
```
CrIDM1: CRO_05G006800
CaMYB123: Caacu.v4.13G012370
MYB123

CrIDM4: CRO_02G002580 
CaMYB5: Caacu.v4.17G000670
MYB005

```{r}
Tip_table <- data.frame(
  label = CAA_CRO_MYB_tree$tip.label
) %>% 
  mutate(label_2 = case_when(
    str_detect(label, "Arabidopsis") ~ str_extract(label, "MYB\\d+"),
    str_detect(label, "CRO") ~ str_remove_all(label, "\\.\\d+$"),
    str_detect(label, "Caacu.v4.") ~ str_remove_all(label, "\\.\\d+$"),
    T ~ ""
  )) %>% 
  mutate(label_3 = case_when(
   # str_detect(label, "05G006800") ~ "CrIDM1",
  #   str_detect(label, "MYB123") ~ "AtTT2",
   #  str_detect(label, "Caacu.v4.13G012370") ~ "CaMYB123",
    str_detect(label, "MYB005") ~ "AtMYB5",
    str_detect(label, "02G002580") ~ "CrIDM4",
    str_detect(label, "Caacu.v4.17G000670") ~ "CaMYB5 (17G000670) ",
    T ~ str_remove_all(label_2, "Caacu.v4.") 
  )) %>% 
  mutate(species = case_when(
    str_detect(label, "Arabidopsis") ~ "A. thaliana",
    str_detect(label, "^CRO") ~ "C. roseus",
    str_detect(label, "^Caac") ~ "C. acuminata", 
    T ~ "other"
  )) %>% 
  mutate(index = 1:n()) 

head(Tip_table)
```
```{r}
CAA_CRO_MYB_tree_nice <- full_join(
  CAA_CRO_MYB_tree, Tip_table, by = "label"
)
```

# Make tree 
## All MYBs 
```{r}
CAA_CRO_MYB_tree_nice %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label_2), size = 2.25) +
  geom_nodelab(aes(subset = as.numeric(label) >= 70, label = label),
               size = 1.8, geom = "text", hjust = 2.2, vjust = -0.75) + 
  scale_color_manual(values = brewer.pal(8, "Set2")[c(1, 2, 4)]) +
  theme(
    legend.position = c(0.1, 0.1),
    legend.text = element_text(face = "italic")
  )

ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree.svg", height = 16, width = 16)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree.pdf", height = 16, width = 16)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree.png", height = 16, width = 16)
```
```{r}
CAA_CRO_MYB_tree_nice %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label_3), size = 4) +
  geom_nodelab(aes(subset = as.numeric(label) >= 70, label = label),
               size = 1.8, geom = "text", hjust = 2.2, vjust = -0.75) + 
  scale_color_manual(values = brewer.pal(8, "Set2")[c(1, 2, 4)]) +
  theme(
    legend.position = c(0.1, 0.1),
    legend.text = element_text(face = "italic")
  )

ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_small.svg", height = 12, width = 12)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_small.pdf", height = 12, width = 12)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_small.png", height = 12, width = 12)
```


## AtTT2/CrIDM1 clade
CrIDM1: CRO_05G006800
CaMYB123: Caacu.v4.13G012370
```{r}
Tip_table %>% 
  filter(str_detect(label, "MYB123|CRO_05G006800|Caacu.v4.13G012370"))
```
```{r}
IDM1_mrca_node <- getMRCA(CAA_CRO_MYB_tree, tip = c("Caacu.v4.13G012370.2", 
                                                    "CRO_05G006800.1",
                                                    "Arabidopsis-thaliana-At5g35550-At2R-MYB123"
                                                    )) 

descendant_tips <- CAA_CRO_MYB_tree$tip.label[offspring(CAA_CRO_MYB_tree, IDM1_mrca_node)]

subset_tree_IDM1 <- treeio::drop.tip(
  CAA_CRO_MYB_tree_nice, 
  CAA_CRO_MYB_tree$tip.label[!(CAA_CRO_MYB_tree$tip.label %in% descendant_tips)]
                        )
```

```{r}
subset_tree_IDM1 %>% 
  ggtree(
    #layout = "circular",
    #branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label_3), size = 3, fontface = "italic") +
  geom_nodelab(aes(
    subset = label >= 70, 
                   label = label),
               size = 3, geom = "text", hjust = 1.2, vjust = -0.75) +
  scale_color_manual(values = c(brewer.pal(8, "Set2")[c(1, 2, 4)])) +
  labs(color = "species") +
  #scale_x_continuous(limits = c(0, 0.4)) +
  theme(
   legend.position = c(0.83, 0.8),
   legend.text = element_text(face = "italic", size = 10),
   title = element_text(size = 10)
  ) 
```
Placed at the base of the tree. 

## AtMYB5/CrIDM4 clade 
CrIDM4: CRO_02G002580 
CaMYB5: Caacu.v4.17G000670
```{r}
Tip_table %>% 
  filter(str_detect(label, "MYB005|CRO_02G002580|Caacu.v4.17G000670|Caacu.v4.2G016000"))
```

```{r}
IDM4_mrca_node <- getMRCA(CAA_CRO_MYB_tree, tip = c("Caacu.v4.17G000670.1", 
                                                    "Caacu.v4.2G016000.1", "CRO_02G002580.1",
                                                    "Arabidopsis-thaliana-At3g13540-At2R-MYB005"
                                                    )) 

descendant_tips <- CAA_CRO_MYB_tree$tip.label[offspring(CAA_CRO_MYB_tree, IDM4_mrca_node)]

subset_tree_IDM4 <- treeio::drop.tip(
  CAA_CRO_MYB_tree_nice, 
  CAA_CRO_MYB_tree$tip.label[!(CAA_CRO_MYB_tree$tip.label %in% descendant_tips)]
                        )
```

```{r}
IDM4_tree <- subset_tree_IDM4 %>% 
  ggtree(
    #layout = "circular",
    #branch.length = "none"
  ) +
  geom_tippoint(aes(color = species), size = 3) +
  geom_tiplab(aes(label = label_3), size = 3, fontface = "italic") +
  geom_nodelab(aes(
    subset = label >= 70, 
                   label = label),
               size = 3, geom = "text", hjust = 1.2, vjust = -0.75) +
  scale_color_manual(values = c(brewer.pal(8, "Set2")[c(1, 2, 4)])) +
  labs(color = "species") +
  scale_x_continuous(limits = c(0, 0.4)) +
  theme(
   legend.position = c(0.83, 0.8),
   legend.text = element_text(face = "italic", size = 10)
  ) 

IDM4_tree

ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_IDM4.svg", height = 3, width = 4)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_IDM4.pdf", height = 3, width = 4)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_IDM4.png", height = 3, width = 4)
```
# bHLHs
## Tree data
```{r}
CAA_CRO_bHLH_tree <- read.tree("../Results/Phylogeny/RAxML_bipartitions.bhlh_out")
CAA_CRO_bHLH_tree
```
* BIS: 08G020510, 08G020500, 08G020490, 02G019960, Caacu.v4.4G008460 
* IDB1: AtBHLH042, 04G007340, Caacu.v4.2G013030, Caacu.v4.2G003530, Caacu.v4.2G003530


```{r}
BHLH_Tip_table <- data.frame(
  label = CAA_CRO_bHLH_tree$tip.label
) %>% 
  mutate(label_2 = case_when(
    str_detect(label, "AtbHLH|^AT") ~ label,
    str_detect(label, "CRO") ~ str_remove_all(label, "\\.\\d+$"),
    str_detect(label, "Caac") ~ str_remove_all(label, "\\.\\d+$"),
    T ~ ""
  )) %>% 
  mutate(label_3 = case_when(
    str_detect(label, "08G020510") ~ "CrBIS2",
    str_detect(label, "08G020500") ~ "CrBIS1",
    str_detect(label, "08G020490") ~ "CrBIS3",
    str_detect(label, "02G019960") ~ "CrBISp",
    str_detect(label, "Caacu.v4.4G008460") ~ "CaBIS",
    str_detect(label, "AtbHLH042") ~ "AtTT8",
    str_detect(label, "04G007340") ~ "CrIDB1",
    #str_detect(label, "Caacu.v4.2G013030") ~ "Caacu.v4.2G013030",
    str_detect(label, "Caacu.v4.2G003530") ~ "2G003530",
    str_detect(label, "Caacu.v4.7G014480") ~ "CaTT8 (7G014480)",
    T ~ ""
  )) %>% 
  mutate(species = case_when(
    str_detect(label, "AtbHLH|^AT") ~ "A. thaliana",
    str_detect(label, "^CRO") ~ "C. roseus",
    str_detect(label, "^Caac") ~ "C. acuminata", 
    T ~ "other"
  )) %>% 
  mutate(index = 1:n()) 

head(BHLH_Tip_table)
```
```{r}
CAA_CRO_bHLH_tree_nice <- full_join(
  CAA_CRO_bHLH_tree, BHLH_Tip_table, by = "label"
)
```

# Make tree
## All bHLHs
```{r}
CAA_CRO_bHLH_tree_nice %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label_2), size = 2.25) +
  geom_nodelab(aes(subset = as.numeric(label) >= 70, label = label),
               size = 1.8, geom = "text", hjust = 2.2, vjust = -0.75) + 
  scale_color_manual(values = brewer.pal(8, "Set2")[c(1, 2, 4)]) +
  theme(
    legend.position = c(0.1, 0.1),
    legend.text = element_text(face = "italic")
  )

ggsave("../Results/R_output/AT_CRO_CAA_bHLH_tree.svg", height = 16, width = 16)
ggsave("../Results/R_output/AT_CRO_CAA_bHLH_tree.pdf", height = 16, width = 16)
ggsave("../Results/R_output/AT_CRO_CAA_bHLH_tree.png", height = 16, width = 16)
```
```{r}
CAA_CRO_bHLH_tree_nice %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label_3), size = 4) +
  geom_nodelab(aes(subset = as.numeric(label) >= 70, label = label),
               size = 1.8, geom = "text", hjust = 2.2, vjust = -0.75) + 
  scale_color_manual(values = brewer.pal(8, "Set2")[c(1, 2, 4)]) +
  theme(
    legend.position = c(0.1, 0.1),
    legend.text = element_text(face = "italic")
  )

ggsave("../Results/R_output/AT_CRO_CAA_bHLH_tree_small.svg", height = 12, width = 12)
ggsave("../Results/R_output/AT_CRO_CAA_bHLH_tree_small.pdf", height = 12, width = 12)
ggsave("../Results/R_output/AT_CRO_CAA_bHLH_tree_small.png", height = 12, width = 12)
```
## BIS clade
* BIS: 08G020510, 08G020500, 08G020490, 02G019960, Caacu.v4.4G008460 

```{r}
BHLH_Tip_table %>% 
  filter(str_detect(label, "08G020510|08G020500|08G020490|02G019960|Caacu.v4.4G008460"))
```
```{r}
BIS_mrca_node <- getMRCA(CAA_CRO_bHLH_tree, tip = c("Caacu.v4.4G008460.4", "CRO_08G020510.1",
                                                    "CRO_02G019960.3", "CRO_08G020490.1", 
                                                    "CRO_08G020500.1")) 

descendant_tips <- CAA_CRO_bHLH_tree$tip.label[offspring(CAA_CRO_bHLH_tree, BIS_mrca_node)]

subset_tree_BIS <- treeio::drop.tip(
  CAA_CRO_bHLH_tree_nice, 
  CAA_CRO_bHLH_tree$tip.label[!(CAA_CRO_bHLH_tree$tip.label %in% descendant_tips)]
                        )
```

```{r}
BIS_tree <- subset_tree_BIS %>% 
  ggtree(
    #layout = "circular",
    #branch.length = "none"
  ) +
  geom_tippoint(aes(color = species), size = 3) +
  geom_tiplab(aes(label = label_3), size = 3, fontface = "italic") +
  geom_nodelab(aes(
    subset = label >= 70, 
                   label = label),
               size = 3, geom = "text", hjust = 1.2, vjust = -0.75) +
  xlim(c(0, 1)) +
  scale_color_manual(values = c(brewer.pal(8, "Set2")[c(2, 4)])) +
  labs(color = "species") +
  # theme_classic() +
  theme(
   legend.position = "bottom",
   legend.text = element_text(face = "italic")
  ) 

BIS_tree

ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_BIS.svg", height = 4, width = 5)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_BIS.pdf", height = 4, width = 5)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_BIS.png", height = 4, width = 5)
```

## TT8/CrIDB1 clade 
* IDB1: AtBHLH042, 04G007340, Caacu.v4.2G013030, Caacu.v4.2G003530, Caacu.v4.2G003530
```{r}
BHLH_Tip_table %>% 
  filter(str_detect(label, "AtbHLH042|04G007340|Caacu.v4.2G013030|Caacu.v4.2G003530|Caacu.v4.7G014480"))
```

```{r}
IDB_mrca_node <- getMRCA(CAA_CRO_bHLH_tree, tip = c("Caacu.v4.2G003530.3", "Caacu.v4.7G014480.1",
                                                    "CRO_04G007340.2", "AtbHLH042")) 
descendant_tips <- CAA_CRO_bHLH_tree$tip.label[offspring(CAA_CRO_bHLH_tree, IDB_mrca_node)]

subset_tree_IDB <- treeio::drop.tip(
  CAA_CRO_bHLH_tree_nice, 
  CAA_CRO_bHLH_tree$tip.label[!(CAA_CRO_bHLH_tree$tip.label %in% descendant_tips)]
                        )
```

```{r}
IDB1_tree <- subset_tree_IDB %>% 
  ggtree(
    #layout = "circular",
    #branch.length = "none"
  ) +
  geom_tippoint(aes(color = species), size = 3) +
  geom_tiplab(aes(label = label_3), size = 3, fontface = "italic") +
  geom_nodelab(aes(
    subset = label >= 70, 
                   label = label),
               size = 3, geom = "text", hjust = 1.2, vjust = -0.75) +
  xlim(c(0, 1.5)) +
  scale_color_manual(values = c(brewer.pal(8, "Set2")[c(1, 2, 4)])) +
  labs(color = "species") +
  #theme_classic() +
  theme(
   legend.position = "bottom",
   legend.text = element_text(face = "italic")
  ) 

IDB1_tree

ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_IDB.svg", height = 4, width = 4)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_IDB.pdf", height = 4, width = 4)
ggsave("../Results/R_output/AT_CRO_CAA_MYB_tree_IDB.png", height = 4, width = 4)
```
# assemble trees
```{r}
wrap_plots(IDM4_tree +
             theme(legend.position = "top"), 
           IDB1_tree + 
             theme(legend.position = "none"), 
           BIS_tree +
             theme(legend.position = "none"), 
           nrow = 3)

ggsave("../Results/R_output/Assembled_TF_trees.svg", height = 6, width = 4)
ggsave("../Results/R_output/Assembled_TF_trees.png", height = 6, width = 4)
```

